Cross correlation function line-profile variations in radial-velocity measurements by fitting a Skew Normal distribution \thanks{Based on observations collected at the La Silla Parana Observatory, ESO (Chile), with the HARPS spectrograph at the 3.6-m telescope.}

### 2018-03-15 manuscript.tex
Added new .tex file for paper

\documentclass{aa}
%\documentclass[referee]{aa}
\usepackage{natbib}
%\usepackage{lscape}
\usepackage{pdflscape} % or {lscape}
\usepackage{wasysym}
\usepackage[varg]{txfonts}
\usepackage{graphicx}
\usepackage{hyperref}
\bibpunct{(}{)}{;}{a}{}{,} % to follow the A&A style
\usepackage{longtable}

\usepackage{afterpage}

\usepackage[usenames,dvipsnames]{color}
\definecolor{mygreen}{rgb}{0,0.5,0}
\definecolor{myorange}{rgb}{0.5,0.5,0}
\definecolor{myred}{rgb}{0.5,0,0}

@@

\newcommand{\jessi}[1]{{\color{Purple}[[\textbf{Jessi: }#1]]}}
\newcommand{\xavier}[1]{{\color{blue}[[\textbf{Xavier: }#1]]}}
\newcommand{\umberto}[1]{{\color{green}[[\textbf{Umberto: }#1]]}}

\begin{document}

\title{Cross correlation function line-profile variations in radial-velocity measurements by fitting a Skew Normal distribution
\thanks{Based on observations collected at the La Silla Parana Observatory,
ESO (Chile), with the HARPS spectrograph at the 3.6-m telescope.}}

@@

\author{U. Simola \inst{1,2}
	    \thanks{\email{umberto.simola@yale.edu}}
	    \and X. Dumusque\inst{3}
	    \thanks{Society in Science -- Branco Weiss Fellow (url: \url{http://www.society-in-science.org})}    
	    \and Jessi Cisewski-Kehe\inst{2}
	    }

\institute{Department of Statistical Sciences, University of Padua, Padua, Italy
	      \and Department of Statistics and Data Science, Yale University, New Haven, CT, USA
	      \and Observatoire de Gen\`eve, Universit\'e de Gen\`eve, 51 ch. des Maillettes, CH-1290 Versoix, Switzerland 
	      }

\date{Received XXX; accepted XXX}

\abstract
% Context, Aims, Methods, Results, Conclu (not mandatory)
{{\bf An exoplanet or stellar companion will induce a variation in the radial velocity (RV) of its host star, which will translate into a pure-Doppler shift of the stellar spectrum and thus a velocity shift of the cross-correlation function (CCF) without changing its shape. The barycenter of the CCF is therefore a measure of the stellar RV. We generally measure this barycenter by fitting a Normal distribution to the CCF and then retreving its first moment, i.e. its mean. However, some astrophysical processes, like oscillations, granulation, stellar activity and single-lined spectroscopic binaries will modify the asymetry of the CCF, which will introduce a spurious radial-velocity variation and perturb the detection of exoplanets. To decorrelate the signals introduced by exoplanets from the one introduced by other astrophysical processes, we use the higher moments of the CCF. Note that in the case of high-precision RV surveys, mainly focusing on non-blended GKM dwarfs, stellar activity is the main perturbing signal.}}
{We propose here to measure the moment of the CCF using a Skew Normal (SN) distribution, which compared to the Normal distribution generally used, provides an extra parameter to model the CCF natural asymmetry.}
{We analyzed 5 stars with different activity levels and different signal-to-noise ratio levels. In each case, we compare the results obtained from the Normal fitting of the CCF and from the SN fitting. We also estimate rigorous errors for the different moments of the CCF using a bootstrap analysis.}
{The correlations between the RVs and the CCF asymmetry or RVs and the CCF width is always stronger when using the parameters derived from the SN. Therefore all the moments of the CCF are more sensitive to activity, which is a huge gain to probe stellar activity. RVs derived from a SN distribution are more sensitive to activity, which is interesting when looking for rotational periods or characterizing better stellar activity. However, once correcting the RVs from stellar activity signanol using a linear combination of the CCF asymmetry and width, the RV residuals obtained from a Normal or SN fitting are very similar. The precision on the asymmetric parameter derived from the SN is also 15\% better than the one derived on the common Bisector Inverse Slope Span (BIS SPAN). Besides all these advantages, the errors on the RVs measured by the SN are 60\% greater. This is not a problem at signal-to-noise ratio levels $>$ 150 because other noise sources dominate in this regime, however we have to be careful when deriving precise RVs for low SNR measurements. We also found that in the case of low SNR measurements, errors measured by bootstrap analysis can be 50\% more precise than errors measured using the derivative of the CCF.}
{We strongly encourage the use of the SN distribution to derive the different moments of the CCF, because the derived moments probe better stellar activity signals than when using a Normal distribution.}

\keywords{techniques: radial velocities -- planetary systems -- stars: activity -- methods: data analysis}

\titlerunning{Fitting a SN distribution to CCF}
\authorrunning{U. Simola, X. Dumusque and J. Cisewski}
\maketitle

%-----------------------------------------------------------------------------------------------------------------------------------------------
\section{Introduction}
\label{intro}

By measuring the radial-velocity of a star, it is possible to detect the stellar wobble induced by the gravitational pull of orbiting planets. This so-called radial-velocity (RV) technique is an indirect method as the planet cannot be seen. This poses some difficulties as the RV technique is not only sensitive to planetary signals, but also to different stellar signals. At the meter-per-second level, RV measurements are affected by oscillations \citep[][]{Arentoft-2008,Kjeldsen-1995,Mayor-2003}, granulation \citep[][]{Meunier-2016,Dumusque-2011a,Lindegren-2003,Del-Moro-2004} and activity signals \citep[][]{Dumusque-2014,Meunier-2011,Saar-2009,Desort-2007,Saar-1997}. Other effects can also introduce RV variations, like single-lined spectroscopic binaries (SB1). High-precision RV surveys tends to avoid targeting any SB1 and to observe GKM dwarfs to minimize the impact of oscillations and granulation. High-precision RV surveys also tend to observe quiet stars in terms of stellar activity, however, as most of the star present solar-like magnetic cycles \citep[][]{Lovis-2011b}, it is not possible to observe the same star at a quiet stage for several continous years.
Stellar activity is therefore the main limitation to the detection of Earth-like planey

While the gravitational pull of an orbiting planets introduces a pure-Doppler

The radial velocity (RV) technique is an indirect method to detect exoplanets.

By measuring the radial velocity induced by the stellar wobble induced by a planet orbiting its host star

%The radial velocity (RV) of a star is defined to be the velocity of the center of mass of the star along our line of sight. This quantity can be derived precisely by measuring the Doppler shift of spectral lines produced in stellar atmospheres. For spectrographs that are not stabilised in temperature and pressure, the iodine technique is used, where the light of the star passes through a iodine cell before getting into the spectrograph to imprint the absorption spectrum of iodine on top of the stellar spectrum (The Hamilton spectrograph \citep{Vogt1987} at Lick Observatory, HIRES \citep{Vogt1994} on the Keck 10-m telescope, the Tull spectrograph  \citep{Tull1995}, the High Resolution Spectrograph HRS  \citep{Tull1998}). In this case, if the spectrograph shifts due to changing atmospheric conditions, the iodine and stellar spectra are shifted in the same way. This leads to some complications when reducing the data because one has to decorrelate the iodine spectrum from the stellar spectrum. For spectrographs that are stabilised, the spectrum of a calibration lamp is recorded close to the stellar spectrum on the CCD, which prevents contamination of the stellar spectrum (CORALIE \citep{Queloz-2000a}, The High Accuracy Radial Velocity Planet Searcher (HARPS) \citep{Mayor-2003}, HARPS- N \citep{Cosentino-2012}, SOPHIE \citep{Bouchy:2013aa}, CARMENES \citep{Quirrenbach:2014aa}). For those instruments, reducing the data is easier as the stellar spectrum is not contaminated with iodine absorption lines.

%For stabilised spectrographs, the RV is derived by first correlating the stellar spectrum with a synthetic \citep[][]{Baranne-1996,Pepe-2002a} or an observed stellar template \citep[][]{Anglada-Escude-2012}, which gives an average line profile, generally called Cross Correlation Function (CCF). Then a Normal distribution is fitted to this average profile to get the mean, namely the RV, and the full width at half maximum (FWHM) of the profile. The CCF technique allows for an averaging out of the RV information of thousands of lines and therefore reach a very high signal-to-noise ratio (SNR), which is essential for a good RV precision.

The convection in external layers of solar type stars is responsible for the granulation pattern than can be seen at high spatial resolution on the surface of the Sun. The differences in flux and velocity between upflows and downflows change the Normal profile of spectral lines that become asymmetric with a "C"-shaped profile \citep[][]{Dravins-1981}. The strength of the asymmetry depends on the velocity of the convection, approximately 300\ms for the Sun, but also on the formation depth of spectral lines \citep{Gray-2009}. Since the CCF is an average of all the spectral lines, some strongly asymmetric and some not, its asymmetry is rather small, which is why a Normal distribution is a proper model to fit the CCF. 
%This small asymmetry modifies however slightly the estimated RVs of the star, reducing the accuracy of the measurement, but if this asymmetry does not vary with time, the precision is kept.

@@

By simple algebra, we can invert those equations to get the DP parameters as a function of the CP ones.
%
\begin{eqnarray} \label{eq:2}
\xi &=& \mu - b\omega\delta, \\
\omega &=& \frac{\sigma}{\sqrt{1 - b^2\,\delta^2}}, \nonumber\\
\alpha &=& \frac{R}{\sqrt{b^2 - (1-b^2)\,R^2}},\,R = sign(\gamma) \left[\frac{2\,|\gamma|}{4-\pi}\right]^{\frac{1}{3}}. \nonumber
\end{eqnarray}
%
The CP allows us to use the usual methods offered by the inferential theory and we will use this parametrisation for the rest of the paper. Further details about DP, CP and statistical properties of the SN are treated in rigorous mathematical and statistical viewpoints in the book by \cite{Azzalini2014}.
%-----------------------------------------------------------------------------------------------------------------------------------------------
\section{Fitting the Skew Normal distribution to the CCF} \label{sec:3}

@@

Finally, for the errors in CCF asymmetry, we see that the boostrap analysis using the Normal fitting gives the same precision as using 2$\times$\emph{noise}, which is the common errors used for BIS SPAN. However, when fitting the SN to the CCF, the asymmetry errors are 15\% smaller. Therefore, the SN fit gives a better precision in CCF asymmetry than what can be reached using BIS SPAN.

@@

The RVs derived using the SN fitting present a higher sensitivity to stellar activity. However, once correcting the RVs from stellar activity using a linear combination of the width and asymmetry parameter of the CCF (FWHM and BIS SPAN in the Normal case, and SN FWHM and $\gamma_1$ in the SN case), the RV residuals in both the Normal and SN case show similar scatter. Therefore, the SN fitting does not improve stellar activity correction when using a simple linear combination with the width and asymmetry of the CCF. However, because the amplitude of the variations is larger in the raw SN RVs, correction techniques based on Gaussian Processes might be more efficient \citep[][]{Haywood-2014,Faria-2016a,Dumusque:2017aa}.

When looking at the correlation between the asymmetry and width parameters of the CCF (FWHM and BIS SPAN or the alternative indicators in \citet{Figueira-2013} in the Normal case, and SN FWHM and $\gamma_1$ in the SN case) with respect to the RVs (RVs in the Normal case or SN RVs in the SN case), we observe that the correlations are always stronger for the parameters of the SN. Therefore, the SN parameters are more sensitive to activity. In the case of Tau Ceti, which is at very low activity level, we find a significant correlation of 0.57 between $\gamma_1$ and SN RV, while for all the other asymmetric parametrisation, BIS SPAN or the alternative indicators in \citet{Figueira-2013}, the correlations are weaker with a maximum of 0.21.

@@

In the case of the errors when using the SN fitting, the precision on the SN RVs is however 60\% worse than for the RVs. This comes from the fact that because the asymmetry is accounted for in the SN and because the SN RVs and $\gamma_1$ are strongly correlated, the errors in $\gamma_1$ propagates to the errors in SN RVs. This is not so important for SNR $>$ 150 for which the errors are below the \ms precision of HARPS and the \ms variation induced by stellar activity signals even in low activity cases \footnote{As we can see in papers publishing HARPS results using an additional jitter to consider stellar activity noise not accounted for in the \emph{noise} parameter, the extra jitter is always greater than 0.8\ms \citep[e.g.][]{Udry:2017aa,Feng:2017aa,Diaz:2016ab}.}. These larger errors in RVs can be a problem for low SNR measurements. In the case of the CCF width, the error derived when using the Normal or SN fitting are very similar. Finally for the asymetry, we see that the error in $\gamma_1$ is $\sim$15\% smaller than the error in BIS SPAN.

@@

%%-----------------------------------------------------------------------------------------------------------------------------------------------
\bibliographystyle{aa}
\bibliography{dumusque_bibliography}

>>>>

\documentclass[11pt, oneside]{article}   
\usepackage{geometry}                		
\geometry{letterpaper}                   		
\usepackage{graphicx}

\usepackage{amssymb}

\usepackage{natbib}
\bibliographystyle{abbrvnat}
\usepackage{aas_macros} % AAS journal abbreviations

\usepackage{lscape}
%\usepackage{pdflscape} % or {lscape}
\usepackage{wasysym}
\usepackage[varg]{txfonts}
\usepackage{graphicx}
\usepackage{hyperref}
\bibpunct{(}{)}{;}{a}{}{,} % to follow the A&A style
\usepackage{longtable}
\usepackage{afterpage}
\usepackage[usenames,dvipsnames]{color}
\definecolor{mygreen}{rgb}{0,0.5,0}
\definecolor{myorange}{rgb}{0.5,0.5,0}
\definecolor{myred}{rgb}{0.5,0,0}

@@

\title{Measuring cross correlation function line-profile variations in radial-velocity measurements via a Skew Normal distribution}
%\thanks{Based on observations collected at the La Silla Parana Observatory,
%ESO (Chile), with the HARPS spectrograph at the 3.6-m telescope.}}

@@

\author{U. Simola
	    \and X. Dumusque
	    \and Jessi Cisewski-Kehe
	    }

%\author{U. Simola \inst{1,2}
%	    \thanks{\email{umberto.simola@yale.edu}}
%	    \and X. Dumusque\inst{3}
%	    \thanks{Society in Science -- Branco Weiss Fellow (url: \url{http://www.society-in-science.org})}    
%	    \and Jessi Cisewski\inst{2}
%	    }
%
%\institute{Department of Statistical Sciences, University of Padua, Padua, Italy
%	      \and Department of Statistics and Data Science, Yale University, New Haven, CT, USA
%	      \and Observatoire de Gen\`eve, Universit\'e de Gen\`eve, 51 ch. des Maillettes, CH-1290 Versoix, Switzerland 
%	      }

\begin{document}
\maketitle

%\abstract
{\bf Abstract}\\
% Context, Aims, Methods, Results, Conclu (not mandatory)
{When working with radial velocities for detecting Extrasolar planets and when using data from stabilized spectrographs,
the different moments of the cross-correlation function (CCF) are used to measure the radial velocity of the star but also the changes in the shape of the CCF. Those changes are due to stellar activity and therefore the best precision on these moments is required to de-correlate exoplanet signals from spurious RV signals originating from stellar activity.}
{We propose here to measure those moments using a Skew Normal (SN) distribution, which compared to the Normal distribution generally used, provides an extra parameter to model the CCF natural asymmetry.}
{We analyzed 5 stars with different activity levels and different signal-to-noise ratio levels. In each case, we compare the results obtained from the Normal fitting of the CCF and from the SN fitting. We also estimate rigorous errors for the different moments of the CCF using a bootstrap analysis.}
{The correlations between the RVs and the CCF asymmetry or RVs and the CCF width is always stronger when using the parameters derived from the SN. Therefore all the moments of the CCF are more sensitive to activity, which is a huge gain to probe stellar activity. RVs derived from a SN distribution are more sensitive to activity, which is interesting when looking for rotational periods or characterizing better stellar activity. However, once correcting the RVs from stellar activity signal using a linear combination of the CCF asymmetry and width, the RV residuals obtained from a Normal or SN fitting are very similar. The precision on the asymmetric parameter derived from the SN is also 15\% better than the one derived on the common Bisector Inverse Slope Span (BIS SPAN). Besides all these advantages, the errors on the RVs measured by the SN are 60\% greater. This is not a problem at signal-to-noise ratio levels $>$ 150 because other noise sources dominate in this regime, however we have to be careful when deriving precise RVs for low SNR measurements. We also found that in the case of low SNR measurements, errors measured by bootstrap analysis can be 50\% more precise than errors measured using the derivative of the CCF.}
{We strongly encourage the use of the SN distribution to derive the different moments of the CCF, because the derived moments probe better stellar activity signals than when using a Normal distribution.}

%\keywords{techniques: radial velocities -- planetary systems -- stars: activity -- methods: data analysis}

\section{Introduction}
\label{intro}

The radial velocity (RV) of a star is defined to be the velocity of the center of mass of the star along our line of sight. This quantity can be derived precisely by measuring the Doppler shift of spectral lines produced in stellar atmospheres. For spectrographs that are not stabilised in temperature and pressure, the iodine technique is used, where the light of the star passes through a iodine cell before getting into the spectrograph to imprint the absorption spectrum of iodine on top of the stellar spectrum (The Hamilton spectrograph \citep{Vogt1987} at Lick Observatory, HIRES \citep{Vogt1994} on the Keck 10-m telescope, the Tull spectrograph  \citep{Tull1995}, the High Resolution Spectrograph HRS  \citep{Tull1998}). In this case, if the spectrograph shifts due to changing atmospheric conditions, the iodine and stellar spectra are shifted in the same way. This leads to some complications when reducing the data because one has to decorrelate the iodine spectrum from the stellar spectrum. For spectrographs that are stabilised, the spectrum of a calibration lamp is recorded close to the stellar spectrum on the CCD, which prevents contamination of the stellar spectrum (CORALIE \citep{Queloz-2000a}, The High Accuracy Radial Velocity Planet Searcher (HARPS) \citep{Mayor-2003}, HARPS- N \citep{Cosentino-2012}, SOPHIE \citep{Bouchy:2013aa}, CARMENES \citep{Quirrenbach:2014aa}). For those instruments, reducing the data is easier as the stellar spectrum is not contaminated with iodine absorption lines.

For stabilized spectrographs, the RV is derived by first correlating the stellar spectrum with a synthetic \citep[][]{Baranne-1996,Pepe-2002a} or an observed stellar template \citep[][]{Anglada-Escude-2012}, which gives an average line profile, generally called Cross Correlation Function (CCF). Then a Normal distribution is fitted to this average profile to get the mean, namely the RV, and the full width at half maximum (FWHM) of the profile. The CCF technique allows for an averaging out of the RV information of thousands of lines and therefore reach a very high signal-to-noise ratio (SNR), which is essential for a good RV precision.

The convection in external layers of solar type stars is responsible for the granulation pattern than can be seen at high spatial resolution on the surface of the Sun. The differences in flux and velocity between upflows and downflows change the Normal profile of spectral lines that become asymmetric with a "C"-shaped profile \citep[][]{Dravins-1981}. The strength of the asymmetry depends on the velocity of the convection, approximatively 300\ms for the Sun, but also on the formation depth of spectral lines \citep{Gray-2009}. Since the CCF is an average of all the spectral lines, some strongly asymmetric and some not, its asymmetry is rather small, which is why a Normal distribution is a proper model to fit the CCF. This small asymmetry modifies however slightly the estimated RVs of the star, reducing the accuracy of the measurement, but if this asymmetry does not vary with time, the precision is kept.

@@

By simple algebra, we can invert those equations to get the DP parameters as a function of the CP ones.
%
\begin{eqnarray} \label{eq:2}
\xi &=& \mu - b\omega\delta, \\
\omega &=& \frac{\sigma}{\sqrt{1 - b^2\,\delta^2}}, \nonumber\\
\alpha &=& \frac{R}{\sqrt{b^2 - (1-b^2)\,R^2}},\,R = sign(\gamma) \left[\frac{2\,|\gamma|}{4-\pi}\right]^{\frac{1}{3}}. \nonumber
\end{eqnarray}
%
The CP allows us to use the usual methods offered by the inferential theory and we will use this parameterization for the rest of the paper. Further details about DP, CP and statistical properties of the SN are treated in rigorous mathematical and statistical viewpoints in the book by \cite{Azzalini2014}.

%-----------------------------------------------------------------------------------------------------------------------------------------------
\section{Fitting the Skew Normal distribution to the CCF} \label{sec:3}

@@

Finally, for the errors in CCF asymmetry, we see that the bootstrap analysis using the Normal fitting gives the same precision as using 2$\times$\emph{noise}, which is the common errors used for BIS SPAN. However, when fitting the SN to the CCF, the asymmetry errors are 15\% smaller. Therefore, the SN fit gives a better precision in CCF asymmetry than what can be reached using BIS SPAN.

@@

The RVs derived using the SN fitting present a higher sensitivity to stellar activity. However, once correcting the RVs from stellar activity using a linear combination of the width and asymmetry parameter of the CCF (FWHM and BIS SPAN in the Normal case, and SN FWHM and $\gamma_1$ in the SN case), the RV residuals in both the Normal and SN case show similar scatter. Therefore, the SN fitting does not improve stellar activity correction when using a simple linear combination with the width and asymmetry of the CCF. However, because the amplitude of the variations is larger in the raw SN RVs, correction techniques based on Gaussian Processes might be more efficient \citep[][]{Haywood-2014,Faria-2016a}.

When looking at the correlation between the asymmetry and width parameters of the CCF (FWHM and BIS SPAN or the alternative indicators in \citet{Figueira-2013} in the Normal case, and SN FWHM and $\gamma_1$ in the SN case) with respect to the RVs (RVs in the Normal case or SN RVs in the SN case), we observe that the correlations are always stronger for the parameters of the SN. Therefore, the SN parameters are more sensitive to activity. In the case of Tau Ceti, which is at very low activity level, we find a significant correlation of 0.57 between $\gamma_1$ and SN RV, while for all the other asymmetric parameterization, BIS SPAN or the alternative indicators in \citet{Figueira-2013}, the correlations are weaker with a maximum of 0.21.

@@

In the case of the errors when using the SN fitting, the precision on the SN RVs is however 60\% worse than for the RVs. This comes from the fact that because the asymmetry is accounted for in the SN and because the SN RVs and $\gamma_1$ are strongly correlated, the errors in $\gamma_1$ propagates to the errors in SN RVs. This is not so important for SNR $>$ 150 for which the errors are below the \ms precision of HARPS and the \ms variation induced by stellar activity signals even in low activity cases \footnote{As we can see in papers publishing HARPS results using an additional jitter to consider stellar activity noise not accounted for in the \emph{noise} parameter, the extra jitter is always greater than 0.8\ms \citep[e.g.][]{Diaz:2016ab}.}. These larger errors in RVs can be a problem for low SNR measurements. In the case of the CCF width, the error derived when using the Normal or SN fitting are very similar. Finally for the asymetry, we see that the error in $\gamma_1$ is $\sim$15\% smaller than the error in BIS SPAN.

@@

%%-----------------------------------------------------------------------------------------------------------------------------------------------
\bibliographystyle{aa}
\bibliography{dumusque_bibliography}
%\bibliography{mybib-SNCCF}

### 2018-03-15 manuscript_umberto.tex
Added new .tex file for paper

\documentclass[11pt, oneside]{article}   
\usepackage{geometry}                		
\geometry{letterpaper}                   		
\usepackage{graphicx}

\usepackage{amssymb}

\usepackage{natbib}
\bibliographystyle{abbrvnat}
\usepackage{aas_macros} % AAS journal abbreviations

\usepackage{lscape}
%\usepackage{pdflscape} % or {lscape}
\usepackage{wasysym}
\usepackage[varg]{txfonts}
\usepackage{graphicx}
\usepackage{hyperref}
\bibpunct{(}{)}{;}{a}{}{,} % to follow the A&A style
\usepackage{longtable}
\usepackage{afterpage}
\usepackage[usenames,dvipsnames]{color}
\definecolor{mygreen}{rgb}{0,0.5,0}
\definecolor{myorange}{rgb}{0.5,0.5,0}
\definecolor{myred}{rgb}{0.5,0,0}

@@

\title{Measuring cross correlation function line-profile variations in radial-velocity measurements via a Skew Normal distribution}
%\thanks{Based on observations collected at the La Silla Parana Observatory,
%ESO (Chile), with the HARPS spectrograph at the 3.6-m telescope.}}

@@

\author{U. Simola
	    \and X. Dumusque
	    \and Jessi Cisewski-Kehe
	    }

%\author{U. Simola \inst{1,2}
%	    \thanks{\email{umberto.simola@yale.edu}}
%	    \and X. Dumusque\inst{3}
%	    \thanks{Society in Science -- Branco Weiss Fellow (url: \url{http://www.society-in-science.org})}    
%	    \and Jessi Cisewski\inst{2}
%	    }
%
%\institute{Department of Statistical Sciences, University of Padua, Padua, Italy
%	      \and Department of Statistics and Data Science, Yale University, New Haven, CT, USA
%	      \and Observatoire de Gen\`eve, Universit\'e de Gen\`eve, 51 ch. des Maillettes, CH-1290 Versoix, Switzerland 
%	      }

\begin{document}
\maketitle

%\abstract
{\bf Abstract}\\
% Context, Aims, Methods, Results, Conclu (not mandatory)
{When working with radial velocities for detecting Extrasolar planets and when using data from stabilized spectrographs,
the different moments of the cross-correlation function (CCF) are used to measure the radial velocity of the star but also the changes in the shape of the CCF. Those changes are due to stellar activity and therefore the best precision on these moments is required to de-correlate exoplanet signals from spurious RV signals originating from stellar activity.}
{We propose here to measure those moments using a Skew Normal (SN) distribution, which compared to the Normal distribution generally used, provides an extra parameter to model the CCF natural asymmetry.}
{We analyzed 5 stars with different activity levels and different signal-to-noise ratio levels. In each case, we compare the results obtained from the Normal fitting of the CCF and from the SN fitting. We also estimate rigorous errors for the different moments of the CCF using a bootstrap analysis.}
{The correlations between the RVs and the CCF asymmetry or RVs and the CCF width is always stronger when using the parameters derived from the SN. Therefore all the moments of the CCF are more sensitive to activity, which is a huge gain to probe stellar activity. RVs derived from a SN distribution are more sensitive to activity, which is interesting when looking for rotational periods or characterizing better stellar activity. However, once correcting the RVs from stellar activity signal using a linear combination of the CCF asymmetry and width, the RV residuals obtained from a Normal or SN fitting are very similar. The precision on the asymmetric parameter derived from the SN is also 15\% better than the one derived on the common Bisector Inverse Slope Span (BIS SPAN). Besides all these advantages, the errors on the RVs measured by the SN are 60\% greater. This is not a problem at signal-to-noise ratio levels $>$ 150 because other noise sources dominate in this regime, however we have to be careful when deriving precise RVs for low SNR measurements. We also found that in the case of low SNR measurements, errors measured by bootstrap analysis can be 50\% more precise than errors measured using the derivative of the CCF.}
{We strongly encourage the use of the SN distribution to derive the different moments of the CCF, because the derived moments probe better stellar activity signals than when using a Normal distribution.}

%\keywords{techniques: radial velocities -- planetary systems -- stars: activity -- methods: data analysis}

\section{Introduction}
\label{intro}

>>>>

\documentclass{aa}
%\documentclass[referee]{aa}
\usepackage{natbib}
\bibliographystyle{abbrvnat}
\usepackage{lscape}
%\usepackage{pdflscape} % or {lscape}
\usepackage{wasysym}
\usepackage[varg]{txfonts}
\usepackage{graphicx}
\usepackage{hyperref}
\bibpunct{(}{)}{;}{a}{}{,} % to follow the A&A style
\usepackage{longtable}
\usepackage{afterpage}
\usepackage[usenames,dvipsnames]{color}
\definecolor{mygreen}{rgb}{0,0.5,0}
\definecolor{myorange}{rgb}{0.5,0.5,0}
\definecolor{myred}{rgb}{0.5,0,0}

@@

\begin{document}

\title{Cross correlation function line-profile variations in radial-velocity measurements by fitting a Skew Normal distribution
\thanks{Based on observations collected at the La Silla Parana Observatory,
ESO (Chile), with the HARPS spectrograph at the 3.6-m telescope.}}

@@

\author{U. Simola \inst{1,2}
	    \thanks{\email{umberto.simola@yale.edu}}
	    \and X. Dumusque\inst{3}
	    \thanks{Society in Science -- Branco Weiss Fellow (url: \url{http://www.society-in-science.org})}    
	    \and Jessi Cisewski\inst{2}
	    }

\institute{Department of Statistical Sciences, University of Padua, Padua, Italy
	      \and Department of Statistics and Data Science, Yale University, New Haven, CT, USA
	      \and Observatoire de Gen\`eve, Universit\'e de Gen\`eve, 51 ch. des Maillettes, CH-1290 Versoix, Switzerland 
	      }

\date{Received XXX; accepted XXX}

\abstract
% Context, Aims, Methods, Results, Conclu (not mandatory)
{When working with radial velocities for detecting Extrasolar planets and when using data from stabilized spectrographs,
the different moments of the cross-correlation function (CCF) are used to measure the radial velocity of the star but also the changes in the shape of the CCF. Those changes are due to stellar activity and therefore the best precision on these moments is required to de-correlate exoplanet signals from spurious RV signals originating from stellar activity.}
{We propose here to measure those moments using a Skew Normal (SN) distribution, which compared to the Normal distribution generally used, provides an extra parameter to model the CCF natural asymmetry.}
{We analyzed 5 stars with different activity levels and different signal-to-noise ratio levels. In each case, we compare the results obtained from the Normal fitting of the CCF and from the SN fitting. We also estimate rigorous errors for the different moments of the CCF using a bootstrap analysis.}
{The correlations between the RVs and the CCF asymmetry or RVs and the CCF width is always stronger when using the parameters derived from the SN. Therefore all the moments of the CCF are more sensitive to activity, which is a huge gain to probe stellar activity. RVs derived from a SN distribution are more sensitive to activity, which is interesting when looking for rotational periods or characterizing better stellar activity. However, once correcting the RVs from stellar activity signal using a linear combination of the CCF asymmetry and width, the RV residuals obtained from a Normal or SN fitting are very similar. The precision on the asymmetric parameter derived from the SN is also 15\% better than the one derived on the common Bisector Inverse Slope Span (BIS SPAN). Besides all these advantages, the errors on the RVs measured by the SN are 60\% greater. This is not a problem at signal-to-noise ratio levels $>$ 150 because other noise sources dominate in this regime, however we have to be careful when deriving precise RVs for low SNR measurements. We also found that in the case of low SNR measurements, errors measured by bootstrap analysis can be 50\% more precise than errors measured using the derivative of the CCF.}
{We strongly encourage the use of the SN distribution to derive the different moments of the CCF, because the derived moments probe better stellar activity signals than when using a Normal distribution.}

\keywords{techniques: radial velocities -- planetary systems -- stars: activity -- methods: data analysis}

\titlerunning{Fitting a SN distribution to CCF}
\authorrunning{U. Simola, X. Dumusque and J. Cisewski}
\maketitle

%-----------------------------------------------------------------------------------------------------------------------------------------------
\section{Introduction}
\label{intro}
